Robust Control of a Fluxonium Qubit

### 2020-11-12 paper/a2unscented.tex
misc updates

\section{Unscented Transformation}
In this section we outline the full unscented sampling procedure.
First, we remark that ALTRO does not natively support complex numbers, so our optimization
is performed under an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$.
\begin{equation}
  H \psi \cong \begin{pmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{pmatrix}
  \begin{pmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{pmatrix}
\end{equation}

@@

The initial $4n + 2d$ sigma points and initial $4n + 2d$ deviant parameters are sampled
from the initial distributions
\begin{equation}
  \begin{bmatrix} \Psi_{0}^{i} \\ \Lambda_{0}^{i} \end{bmatrix} =
  \begin{bmatrix} \bar{\Psi}_{0} \\ 0\end{bmatrix}
    \pm \beta \sqrt{\begin{bmatrix} P_{0} & 0\\ 0 & L_{0}\end{bmatrix}}^{\; i}
\end{equation}
We have written $\bar{\Psi}_{0} = \psi_{0}$. $\beta$
is a hyperparameter which controls the spacing of the covariance contour.
The $\pm$ is understood to take $+$ for $ i \in \{1, \dots, 2n + d\}$ and $-$ for
$i \in \{2n + d + 1, \dots, 4n + 2d\}$. We use the Cholesky factorization
to compute the square root of the
joint covariance matrix, though other methods
such as the principal square root may be employed.
The superscript indicates the $i$\textsuperscript{th} column of the
lower triangular Cholesky factor.
Then, the sigma points are normalized
\begin{equation}
  \Psi_{0}^{i} \gets \frac{\Psi_{0}^{i}}{\sqrt{\Psi_{0}^{i^{T}} \Psi_{0}^{i}}}
\end{equation}
The sigma points are propagated to the next knot point
\begin{equation}
  \Psi_{1} = f(\Psi_{0}, \Lambda_{0})
\end{equation}
The mean and covariance of the sigma points is computed
\begin{align}
  \bar{\Psi}_{1} &= \frac{1}{4n + 2d} \sum_{i = 1}^{4n + 2d} \Psi_{1}^{i}\\
  P_{1} &= \frac{1}{\beta^{2}(4n + 2d)} \sum_{i = 1}^{4n + 2d}
  (\Psi^{i}_{1} - \bar{\Psi}_{1})^{T}(\Psi^{i}_{1} - \bar{\Psi}_{1})
\end{align}
The sigma points are then resampled and propagated to the next knot point using eq B2, B3, and B4.

>>>>

\section{Unscented Transformation}
\label{appendix:unscented}
In this section we outline the full unscented sampling procedure.
First, we remark that ALTRO does not natively support complex numbers, so our optimization
is performed under an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
\cite{leung2017speedup}
\begin{equation}
  H \psi \cong \begin{pmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{pmatrix}
  \begin{pmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{pmatrix}
\end{equation}

@@

The initial $4n + 2d$ sigma points and initial $4n + 2d$ deviant parameters are sampled
from the initial distributions
\begin{equation}\label{eq:uupdate}
  \begin{bmatrix} \Psi_{0}^{i} \\ \Lambda_{0}^{i} \end{bmatrix} =
  \begin{bmatrix} \bar{\Psi}_{0} \\ 0\end{bmatrix}
    \pm \beta \sqrt{\begin{bmatrix} P_{0} & 0\\ 0 & L_{0}\end{bmatrix}}^{\; i}
\end{equation}
We have written $\bar{\Psi}_{0} = \psi_{0}$. $\beta$
is a hyperparameter which controls the spacing of the covariance contour.
The $(\pm)$ is understood to take $(+)$ for $ i \in \{1, \dots, 2n + d\}$ and $(-)$ for
$i \in \{2n + d + 1, \dots, 4n + 2d\}$. We use the Cholesky factorization
to compute the square root of the
joint covariance matrix, though other methods
such as the principal square root may be employed.
The superscript on the matrix square root indicates the $i$\textsuperscript{th}
column (mod $2n + d$) of the lower triangular Cholesky factor.
Then, the sigma points are normalized
\begin{equation}\label{eq:unormalize}
  \Psi_{0}^{i} \gets \frac{\Psi_{0}^{i}}{\sqrt{\Psi_{0}^{i^{T}} \Psi_{0}^{i}}}
\end{equation}
The sigma points are propagated to the next knot point
\begin{equation}\label{eq:upropagate}
  \Psi_{1} = f(\Psi_{0}, \Lambda_{0})
\end{equation}
The mean and covariance of the sigma points are computed
\begin{align}
  \bar{\Psi}_{1} &= \frac{1}{4n + 2d} \sum_{i = 1}^{4n + 2d} \Psi_{1}^{i}\\
  P_{1} &= \frac{1}{\beta^{2}(4n + 2d)} \sum_{i = 1}^{4n + 2d}
  (\Psi^{i}_{1} - \bar{\Psi}_{1})^{T}(\Psi^{i}_{1} - \bar{\Psi}_{1})
\end{align}
The sigma points are then resampled and propagated to the next knot point using
\eqref{eq:uupdate}, \eqref{eq:unormalize}, and \eqref{eq:upropagate}.

### 2020-11-12 paper/a3derivative.tex
misc updates

\section{Derivative Method}
We believe the 1\textsuperscript{st}-order derivative method tends
to outperform the 2\textsuperscript{nd}-order
derivative method in the small, static detuning regime, see Figure 2b,
because second-order contributions to the gate error are small.
The norm of the state derivatives for the 1\textsuperscript{st}- and 2\textsuperscript{nd}-order
derivative methods are provided in Table 2 for the solutions of the optimization outlined
in section V at the gate duration $t_{N} = 38$ns.

%%TODO: the values in this table are not up to date until this comment is removed
\begin{table}[h]
  \begin{tabular}{c | c | c}
    Method & ${\lvert \braket{\partial_{f_{q}} \psi_{N} | \partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
    & ${\lvert \braket{\partial^{2}_{f_{q}} \psi_{N} | \partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
    \hline
    D-1 & 2.051 & 17.935\\
    D-2 & 2.511 & 7.628\\
  \end{tabular}
  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
    for $Z/2$ gates optimized using the derivative methods. The norms are computed
    at the end of the gate's duration $t_{N} = 38$ns.}
\end{table}

The dynamics for the derivative methods can be integrated efficiently using exponential integrators.
General exponential integrators break the dynamics into a linear term and a non-linear term. For example, consider
integrating the dynamics of the first state derivative $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
H \ket{\partial_{\lambda} \psi} + (\partial_{\lambda} H) \ket{\psi}$ in units of $i\hbar = 1$.
The linear term is $L = H$ and the non-linear term is $N = (\partial_{\lambda} H) \ket{\psi}$.
With zero-order hold on the controls the exact propagation is
\begin{equation}
  \begin{aligned}
    \ket{\partial_{\lambda} \psi_{k + 1}} &= \exp(L_{k} \Delta t_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
    &+ \int_{0}^{\Delta t_{k}} \exp(L_{k}(\Delta t_{k} - t^{'})) N(t_{k} + t^{'}) dt^{'}\\
  \end{aligned}
\end{equation}
General exponential integrators proceed by breaking the integral into a discrete sum, similar to the procedure
for Runge-Kutta schemes. We use a simple approximation known as the Lawson-Euler method \cite{berland2005solving}
\begin{equation}
  \begin{aligned}
    \ket{\partial_{\lambda} \psi_{k + 1}} &\approx \exp(L_{k} \Delta t_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
    &+ \exp(L_{k} \Delta t_{k}) N_{k}\\
  \end{aligned}
\end{equation}
This method provides a good tradeoff between accuracy and efficiency, requiring one unique matrix
exponential computation per stage. Exact integration accuracy is not of the utmost importance because the
state derivatives guide the robust optimization, and do not correspond to experimental parameters
which must be realized with high accuracy.

>>>>

\section{Derivative Method}
\label{appendix:derivative}
We comment on the relative performance of the derivative methods and then
outline how to efficiently integrate their dynamics.
We believe the 1\textsuperscript{st}-order derivative method tends
to outperform the 2\textsuperscript{nd}-order
derivative method in the small, static detuning regime, see Figure 2b,
because second-order contributions to the gate error are small.
The norm of the state derivatives for the 1\textsuperscript{st}- and 2\textsuperscript{nd}-order
derivative methods are provided in Table \ref{tab:dnorm} for the solutions of the optimization outlined
in section V at the gate duration $t_{N} = 38$ns.

%%TODO: the values in this table are not up to date until this comment is removed
\begin{table}[h]
  \label{tab:dnorm}
  \begin{tabular}{c | c | c}
    Method & ${\lvert \braket{\partial_{f_{q}} \psi_{N} | \partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
    & ${\lvert \braket{\partial^{2}_{f_{q}} \psi_{N} | \partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
    \hline
    D-1 & 2.051 & 17.935\\
    D-2 & 2.511 & 7.628\\
  \end{tabular}
  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
    for $Z/2$ gates optimized using the derivative methods. The norms are computed
    at the end of the gate's duration $t_{N} = 38$ns.}
\end{table}

The dynamics for the derivative methods can be integrated efficiently using exponential integrators
\cite{berland2005solving, einkemmer2017performance}.
General exponential integrators break the dynamics into a linear term and a non-linear term. For example, consider
integrating the dynamics of the first state derivative $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
H \ket{\partial_{\lambda} \psi} + (\partial_{\lambda} H) \ket{\psi}$ in units of $i\hbar = 1$.
The linear term is $L = H$ and the non-linear term is $N = (\partial_{\lambda} H) \ket{\psi}$.
With zero-order hold on the controls the exact propagation is
\begin{equation}
  \label{eq:dgeneralexp}
  \begin{aligned}
    \ket{\partial_{\lambda} \psi_{k + 1}} &= \exp(\Delta t_{k} L_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
    &+ \int_{0}^{\Delta t_{k}} \exp((\Delta t_{k} - t^{'})L_{k}) N(t_{k} + t^{'}) dt^{'}\\
  \end{aligned}
\end{equation}
General exponential integrators proceed by breaking the integral in \eqref{eq:dgeneralexp}
into a discrete sum, similar to the procedure
for Runge-Kutta schemes. We use a simple approximation known as the Lawson-Euler
method \cite{berland2005solving}
\begin{equation}
  \begin{aligned}
    \ket{\partial_{\lambda} \psi_{k + 1}} &\approx \exp(\Delta t_{k}L_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
    &+ \exp(\Delta t_{k}L_{k}) N_{k}\\
  \end{aligned}
\end{equation}
This method provides a good tradeoff between accuracy and efficiency, requiring one unique matrix
exponential computation per stage. Integration accuracy is not of the utmost importance because the
state derivatives guide the optimization, and do not correspond to experimental parameters
which must be realized with high accuracy.

### 2020-11-12 paper/main.tex
updated many, added pdf
misc updates

%% TITLE
\title{Robust Control of a Fluxonium Qubit}

\author{Thomas Propson}
\email{tcpropson@uchicago.edu}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\author{Brian Jackson}
\author{Zac Manchester}
\affiliation{
  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
}
\author{David I. Schuster}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
}

\date{\today}

\input{0abstract.tex}

\maketitle

@@

\appendix
\input{a1experiment.tex}
\input{a2dissipation.tex}
\input{a3derivative.tex}
\input{a4ctensor.tex}

>>>>

\input{0abstract.tex}
\maketitle

@@

\appendix
\input{a1longitude.tex}
\input{a2unscented.tex}
\input{a3derivative.tex}
